---
title: "DS311 - Basic R Lab Exercise"
author: "Your Name"
date: "8/23/2022"
output:
  pdf_document: default
  html_document: default
subtitle: R Lab Exercise
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# OUTPUT
<!-- # > knitr::opts_chunk$set(echo = TRUE) -->

# Basic R Exercise

## Section 1 - Data Type

**Key Functions**
- typeof()
- as.numeric()
- as.charater()

### Numeric

```{r}
# Numeric - Double precision by default

n1 <- 15  
n1
typeof(n1)

n2 <- 1.5
n2
typeof(n2)
```
# OUTPUT
<!-- # > # Numeric - Double precision by default
# > 
# > n1 <- 15  
# > n1
# [1] 15
# typeof(n1)

# n2 <- 1.5
# > typeof(n1)
# [1] "double"
# > 
# > n2 <- 1.5 -->


### Character

```{r}
# Character

c1 <- "c"
c1
typeof(c1)

c2 <- "a string of text"
c2
typeof(c2)

```
# OUTPUT
<!-- # > # Character
# > 
# > c1 <- "c"
# c1
# typeof(c1)
# > c1
# [1] "c"
# > typeof(c1)
# [1] "character"
# > 
# > c2 <- "a string of text"
# > c2
# [1] "a string of text"
# > typeof(c2)
# [1] "character" -->


### Logical

```{r}
# Logical

l1 <- TRUE
l1
typeof(l1)

l2 <- F
l2
typeof(l2)
```
# OUTPUT
<!-- # Logical
# > 
# > l1 <- TRUE
# l1
# > l1
# [1] TRUE
# > typeof(l1)
# [1] "logical" -->

### Transforming Numerics and Characters

``` {r}
# Transforming numeric into characters
num <- 10
numToChar <- as.character(num)
paste("num Type: ", typeof(num), " | numToChar: ", typeof(numToChar))

# Transforming characters into numeric
char <- "10"
charToNum <- as.numeric(char)
paste("char Type: ", typeof(char), " | charToNum: ", typeof(charToNum))
```
# OUTPUT
<!-- # > # Transforming numeric into characters
# > num <- 10
# > numToChar <- as.character(num)
# > paste("num Type: ", typeof(num), " | numToChar: ", typeof(numToChar))
# [1] "num Type:  double  | numToChar:  character"

# > 
# > # Transforming characters into numeric
# > char <- "10"
# charToNum <- as.numeric(char)
# paste("char Type: ", typeof(char), " | charToNum: ", typeof(charToNum))
# > charToNum <- as.numeric(char)
# > paste("char Type: ", typeof(char), " | charToNum: ", typeof(charToNum))
# [1] "char Type:  character  | charToNum:  double" -->

### Challenge:

Complete the following tasks:

``` {r}
# Check the data type of the following variables
a <- as.integer(500)
b <- as.double(500)
c <- as.character(500)

# Enter your code here!
paste("a: ", a, " | TYPE: ", typeof(a))
paste("b: ", b, " | TYPE: ", typeof(b))
paste("c: ", c, " | TYPE: ", typeof(c))


# Check the data type of the following variable
d <- a / b

# Enter your code here!
paste("d: ", d, " | TYPE: ", typeof(d))


```
# OUTPUT
<!-- > # Check the data type of the following variables
> a <- as.integer(500)
b <- as.double(500)
> b <- as.double(500)
> c <- as.character(500)

# Enter your code here!
> 
> # Enter your code here!
> paste("a: ", a, " | TYPE: ", typeof(a))
[1] "a:  500  | TYPE:  integer"
paste("b: ", b, " | TYPE: ", typeof(b))
paste("c: ", c, " | TYPE: ", typeof(c))


# Check the data type of the following variable
> paste("b: ", b, " | TYPE: ", typeof(b))
[1] "b:  500  | TYPE:  double"
paste("c: ", c, " | TYPE: ", typeof(c))


# Check the data type of the following variable
d <- a / b
> paste("c: ", c, " | TYPE: ", typeof(c))
[1] "c:  500  | TYPE:  character"


# Check the data type of the following variable
d <- a / b

# Enter your code here!
paste("d: ", d, " | TYPE: ", typeof(d))
> 
> 
> # Check the data type of the following variable
> d <- a / b
> 
> # Enter your code here!
> paste("d: ", d, " | TYPE: ", typeof(d))
[1] "d:  1  | TYPE:  double" -->
***

## Section 2 - Data Structure
- is.vector()
- is.matrix
- cbind()
- as.data.frame()

### Vector

```{r}
# Vector

v1 <- c(1, 2, 3, 4, 5)
v1
is.vector(v1)

v2 <- c("a", "b", "c")
v2
is.vector(v2)

v3 <- c(TRUE, TRUE, FALSE, FALSE, TRUE)
v3
is.vector(v3)
```
# OUTPUT
<!-- # Vector
> 
> v1 <- c(1, 2, 3, 4, 5)
v1
> v1
[1] 1 2 3 4 5
> is.vector(v1)
[1] TRUE

v2 <- c("a", "b", "c")
> 
> v2 <- c("a", "b", "c")
> v2
[1] "a" "b" "c"
> is.vector(v2)
[1] TRUE -->

### Matrix

```{r}
# Matrix

m1 <- matrix(c(T, T, F, F, T, F), nrow = 2)
m1
is.matrix(m1)

m2 <- matrix(c("a", "b", 
               "c", "d"), 
               nrow = 2,
               byrow = T)
m2
is.matrix(m2)
```
# OUTPUT
<!-- > # Matrix
> 
> m1 <- matrix(c(T, T, F, F, T, F), nrow = 2)
> m1
     [,1]  [,2]  [,3]
[1,] TRUE FALSE  TRUE
[2,] TRUE FALSE FALSE
> is.matrix(m1)
[1] TRUE
> 
> m2 <- matrix(c("a", "b", 
+                "c", "d"), 
+                nrow = 2,
+                byrow = T)
> m2
     [,1] [,2]
[1,] "a"  "b" 
[2,] "c"  "d" 
> is.matrix(m2)
[1] TRUE -->

### Challenge:

1. Create a vector of the 26 alphabet lower case letters in sequence. 
2. Create a 2 by 13 matrix for the 26 English upper case letter in sequence.

Hint: Check out the "letters" and "LETTERS" key words in R.

```{r}
# Enter your code here.
matrix1 <- matrix()
is.matrix(matrix1)
vector <- c()
chr <- function(n) { rawToChar(as.raw(n)) }
for (x in 97:122)
  vector <- append(vector, chr(x))
print(vector)
matrix1<-matrix(toupper(vector),nrow=2, byrow = T)
print(matrix1)
```
# OUTPUT 
<!-- > # Enter your code here.
> matrix1 <- matrix()
is.matrix(matrix1)
vector <- c()
chr <- function(n) { rawToChar(as.raw(n)) }
for (x in 97:122)
  vector <- append(vector, chr(x))
print(vector)
matrix1<-matrix(toupper(vector),nrow=2, byrow = T)
print(matrix1)
> is.matrix(matrix1)
[1] TRUE
> vector <- c()
> chr <- function(n) { rawToChar(as.raw(n)) }
> for (x in 97:122)
+   vector <- append(vector, chr(x))
> print(vector)
 [1] "a" "b" "c" "d" "e" "f" "g" "h" "i" "j" "k" "l" "m" "n" "o" "p" "q" "r" "s"
[20] "t" "u" "v" "w" "x" "y" "z"
> matrix1<-matrix(toupper(vector),nrow=2, byrow = T)
> print(matrix1)
     [,1] [,2] [,3] [,4] [,5] [,6] [,7] [,8] [,9] [,10] [,11] [,12] [,13]
[1,] "A"  "B"  "C"  "D"  "E"  "F"  "G"  "H"  "I"  "J"   "K"   "L"   "M"  
[2,] "N"  "O"  "P"  "Q"  "R"  "S"  "T"  "U"  "V"  "W"   "X"   "Y"   "Z"  -->

### DataFrame

```{r}
# Data Frame

# Can combine vectors of the same length
vNumeric   <- c(1, 2, 3)
vCharacter <- c("a", "b", "c")
vLogical   <- c(T, F, T)

df1 <- cbind(vNumeric, vCharacter, vLogical)
df1  # Coerces all values to most basic data type

df2 <- as.data.frame(cbind(vNumeric, vCharacter, vLogical))
df2  # Makes a data frame with three different data types
```
# OUTPUT
<!-- > # Data Frame
> 
> # Can combine vectors of the same length
> vNumeric   <- c(1, 2, 3)
> vCharacter <- c("a", "b", "c")
> vLogical   <- c(T, F, T)
> 
> df1 <- cbind(vNumeric, vCharacter, vLogical)
> df1  # Coerces all values to most basic data type
     vNumeric vCharacter vLogical
[1,] "1"      "a"        "TRUE"  
[2,] "2"      "b"        "FALSE" 
[3,] "3"      "c"        "TRUE"  

df2 <- as.data.frame(cbind(vNumeric, vCharacter, vLogical))
df2  # Makes a data frame with three different data types
> 
> df2 <- as.data.frame(cbind(vNumeric, vCharacter, vLogical))
> df2  # Makes a data frame with three different data types
  vNumeric vCharacter vLogical
1        1          a     TRUE
2        2          b    FALSE
3        3          c     TRUE -->

***

## Section 3 - Setup Working Directory and Installing Packages

**Key Functions:**
- getwd()
- setwd()
- install.packages()
- library()

### Setting up your working directory

``` {r}
# Check your current working directory
initial <- "/Users/robert/sfsu/ds311"    # Just to be sure that I will go back in my initial directory
setwd(initial)
wd1 <- getwd()
paste("Current Working Directory: ", wd1)

# Setting the working directory for a project
p <- paste(wd1, "/Hello-World", sep="")
setwd(p)
wd2 <- getwd()
paste("Current Working Directory: ", wd2)
setwd(initial)
```
# OUTPUT
<!-- > # Check your current working directory
> initial <- "/Users/robert/sfsu/ds311"    # Just to be sure that I will go back in my initial directory
> setwd(initial)
> wd1 <- getwd()
> paste("Current Working Directory: ", wd1)
[1] "Current Working Directory:  /Users/robert/sfsu/ds311"
> 
> # Setting the working directory for a project
> p <- paste(wd1, "/Hello-World", sep="")
> setwd(p)
> wd2 <- getwd()
> paste("Current Working Directory: ", wd2)
[1] "Current Working Directory:  /Users/robert/sfsu/ds311/Hello-World"
> setwd(initial) -->

### Installing and Loading Packages

```{r, include=FALSE}
# Install a new package, note the quotation marks
install.packages("mass")

# Install multiple packages at once
install.packages(c("dplyr", "ggplot2"))

# Loading the package, note no quotation marks
library(dplyr)

# Checking the package version
packageVersion("dplyr")

# List all functions in a package
ls("package:ggplot2")

# Loading a function from package
ggplot2::geom_line

# Update all packages
update.packages()

# Unload a package
detach(package:ggplot2, unload=TRUE)

# Help function
help(dplyr)

# Checking the session info
sessionInfo()
```
# OUTPUT
<!-- > # Install a new package, note the quotation marks
> install.packages("mass")
--- Please select a CRAN mirror for use in this session ---

# Install multiple packages at once
install.packages(c("dplyr", "ggplot2"))

# Loading the package, note no quotation marks
library(dplyr)

# Checking the package version
packageVersion("dplyr")

# List all functions in a package
ls("package:g
Secure CRAN mirrors 

 1: 0-Cloud [https]
 2: Australia (Canberra) [https]
 3: Australia (Melbourne 1) [https]
 4: Australia (Melbourne 2) [https]
 5: Australia (Perth) [https]
 6: Austria [https]
 7: Belgium (Brussels) [https]
 8: Brazil (PR) [https]
 9: Brazil (RJ) [https]
10: Brazil (SP 1) [https]
11: Brazil (SP 2) [https]
12: Bulgaria [https]
13: Canada (MB) [https]
14: Canada (ON 3) [https]
15: Chile (Santiago) [https]
16: China (Beijing 2) [https]
17: China (Beijing 3) [https]
18: China (Hefei) [https]
19: China (Hong Kong) [https]
20: China (Guangzhou) [https]
21: China (Lanzhou) [https]
22: China (Nanjing) [https]
23: China (Shanghai 2) [https]
24: China (Shenzhen) [https]
25: Colombia (Cali) [https]
26: Costa Rica [https]
27: Cyprus [https]
28: Czech Republic [https]
29: Denmark [https]
30: East Asia [https]
31: Ecuador (Cuenca) [https]
32: Ecuador (Quito) [https]
33: France (Lyon 1) [https]
34: France (Lyon 2) [https]
35: France (Marseille) [https]
36: France (Paris 1) [https]
37: Germany (Erlangen) [https]
38: Germany (Leipzig) [https]
39: Germany (Göttingen) [https]
40: Germany (Münster) [https]
41: Germany (Regensburg) [https]
42: Greece [https]
43: Hungary [https]
44: Iceland [https]
45: India [https]
46: Indonesia (Banda Aceh) [https]
47: Iran (Mashhad) [https]
48: Italy (Milano) [https]
49: Italy (Padua) [https]
50: Japan (Tokyo) [https]
51: Japan (Yonezawa) [https]
52: Korea (Gyeongsan-si) [https]
53: Korea (Seoul 1) [https]
54: Korea (Ulsan) [https]
55: Malaysia [https]
56: Mexico (Mexico City) [https]
57: Mexico (Texcoco) [https]
58: Morocco [https]
59: Netherlands (Dronten) [https]
60: New Zealand [https]
61: Norway [https]
62: South Africa (Johannesburg) [https]
63: Spain (A Coruña) [https]
64: Spain (Madrid) [https]
65: Sweden (Borås) [https]
66: Sweden (Umeå) [https]
67: Switzerland (Zurich 1) [https]
68: Taiwan (Taipei) [https]
69: Turkey (Denizli) [https]
70: Turkey (Istanbul) [https]
71: Turkey (Mersin) [https]
72: UK (Bristol) [https]
73: UK (London 1) [https]
74: USA (IA) [https]
75: USA (MI) [https]
76: USA (MO) [https]
77: USA (OH) [https]
78: USA (OR) [https]
79: USA (TN) [https]
80: USA (TX 1) [https]
81: Uruguay [https]
82: (other mirrors)

Selection: 
Enter an item from the menu, or 0 to exit
Selection: # Install multiple packages at once
Enter an item from the menu, or 0 to exit
Selection: install.packages(c("dplyr", "ggplot2"))
Enter an item from the menu, or 0 to exit
Selection: 
Enter an item from the menu, or 0 to exit
Selection: # Loading the package, note no quotation marks
Enter an item from the menu, or 0 to exit
Selection: library(dplyr)
Enter an item from the menu, or 0 to exit
Selection: 
Enter an item from the menu, or 0 to exit
Selection: # Checking the package version
Enter an item from the menu, or 0 to exit
Selection: packageVersion("dplyr")
Enter an item from the menu, or 0 to exit
Selection: 
Enter an item from the menu, or 0 to exit
Selection: # List all functions in a package
Enter an item from the menu, or 0 to exit
Selection: ls("package:g
Enter an item from the menu, or 0 to exit
Selection: -->

***

## Section 4 - Problem Solving

Write the code that accomplish the following tasks:

Part a: Assign 4 to variable x

Part b: Assign 12 to variable y

Part c: Print both x and y to check their values

Part d: Divide y by x and assign it to variable z

part e: Print a statement to report your answer in Part d.

Once you finished and knit the RMarkdown file into html file, you should be able to see the message "Congratulation!!  You completed the first exercise in this section!!" in the html document.

```{r}
# Write your code here!
# Part a
x <- 4

# Part b
y <- 12

# Part c
print(paste(x, sep = "   ", y))

# Part d
z <- y/x

# Part e
print(paste("y divided by x is equal to ", z))


# Do not need to change the following code!
if (exists("x") == TRUE | exists("y") == TRUE | exists("z") == TRUE){
  if (x == 4 & y == 12 & z == 3) {
  print("Congratulation!!  You completed the first activity in this class!!")
  } else {
    print("Sorry, you got it wrong!")
  }
} else {
  print("You did not complete the last problem!")
}
```
# OUTPUT
<!-- > # Write your code here!
> # Part a
> x <- 4
> 
> # Part b
> y <- 12
> 
> # Part c
> print(paste(x, sep = "   ", y))
[1] "4   12"
> 
> # Part d
> z <- y/x
> 
> # Part e
> print(paste("y divided by x is equal to ", z))
[1] "y divided by x is equal to  3"
> 
> 
> # Do not need to change the following code!
> if (exists("x") == TRUE | exists("y") == TRUE | exists("z") == TRUE){
+   if (x == 4 & y == 12 & z == 3) {
+   print("Congratulation!!  You completed the first activity in this class!!")
+   } else {
+     print("Sorry, you got it wrong!")
+   }
+ } else {
+   print("You did not complete the last problem!")
+ }
[1] "Congratulation!!  You completed the first activity in this class!!" -->
