---
title: "DS311 - Basic R Lab Exercise"
author: "Robert Harakaly"
date: "8/23/2022"
output:
  pdf_document: default
  html_document: default
subtitle: "R Lab Exercise"
---


## 1

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
df <- read.csv(file = './EDA_Python/data/ames.csv')

```

```{r}
head(df)
nb_row <- nrow(df)
nb_col <- ncol(df)
if (nb_row == 1460)
  print("Good")
if (nb_col == 81)
  print("GOOD2")

```

Id MSSubClass MSZoning LotFrontage LotArea Street Alley LotShape LandContour
1  1         60       RL          65    8450   Pave  <NA>      Reg         Lvl
2  2         20       RL          80    9600   Pave  <NA>      Reg         Lvl
3  3         60       RL          68   11250   Pave  <NA>      IR1         Lvl
4  4         70       RL          60    9550   Pave  <NA>      IR1         Lvl
5  5         60       RL          84   14260   Pave  <NA>      IR1         Lvl
6  6         50       RL          85   14115   Pave  <NA>      IR1         Lvl
  Utilities LotConfig LandSlope Neighborhood Condition1 Condition2 BldgType
1    AllPub    Inside       Gtl      CollgCr       Norm       Norm     1Fam
2    AllPub       FR2       Gtl      Veenker      Feedr       Norm     1Fam
3    AllPub    Inside       Gtl      CollgCr       Norm       Norm     1Fam
4    AllPub    Corner       Gtl      Crawfor       Norm       Norm     1Fam
5    AllPub       FR2       Gtl      NoRidge       Norm       Norm     1Fam
6    AllPub    Inside       Gtl      Mitchel       Norm       Norm     1Fam
  HouseStyle OverallQual OverallCond YearBuilt YearRemodAdd RoofStyle RoofMatl
1     2Story           7           5      2003         2003     Gable  CompShg
2     1Story           6           8      1976         1976     Gable  CompShg
3     2Story           7           5      2001         2002     Gable  CompShg
4     2Story           7           5      1915         1970     Gable  CompShg
5     2Story           8           5      2000         2000     Gable  CompShg
6     1.5Fin           5           5      1993         1995     Gable  CompShg
  Exterior1st Exterior2nd MasVnrType MasVnrArea ExterQual ExterCond Foundation
1     VinylSd     VinylSd    BrkFace        196        Gd        TA      PConc
2     MetalSd     MetalSd       None          0        TA        TA     CBlock
3     VinylSd     VinylSd    BrkFace        162        Gd        TA      PConc
4     Wd Sdng     Wd Shng       None          0        TA        TA     BrkTil
5     VinylSd     VinylSd    BrkFace        350        Gd        TA      PConc
6     VinylSd     VinylSd       None          0        TA        TA       Wood
  BsmtQual BsmtCond BsmtExposure BsmtFinType1 BsmtFinSF1 BsmtFinType2
1       Gd       TA           No          GLQ        706          Unf
2       Gd       TA           Gd          ALQ        978          Unf
3       Gd       TA           Mn          GLQ        486          Unf
4       TA       Gd           No          ALQ        216          Unf
5       Gd       TA           Av          GLQ        655          Unf
6       Gd       TA           No          GLQ        732          Unf
  BsmtFinSF2 BsmtUnfSF TotalBsmtSF Heating HeatingQC CentralAir Electrical
1          0       150         856    GasA        Ex          Y      SBrkr
2          0       284        1262    GasA        Ex          Y      SBrkr
3          0       434         920    GasA        Ex          Y      SBrkr
4          0       540         756    GasA        Gd          Y      SBrkr
5          0       490        1145    GasA        Ex          Y      SBrkr
6          0        64         796    GasA        Ex          Y      SBrkr
  X1stFlrSF X2ndFlrSF LowQualFinSF GrLivArea BsmtFullBath BsmtHalfBath FullBath
1       856       854            0      1710            1            0        2
2      1262         0            0      1262            0            1        2
3       920       866            0      1786            1            0        2
4       961       756            0      1717            1            0        1
5      1145      1053            0      2198            1            0        2
6       796       566            0      1362            1            0        1
  HalfBath BedroomAbvGr KitchenAbvGr KitchenQual TotRmsAbvGrd Functional
1        1            3            1          Gd            8        Typ
2        0            3            1          TA            6        Typ
3        1            3            1          Gd            6        Typ
4        0            3            1          Gd            7        Typ
5        1            4            1          Gd            9        Typ
6        1            1            1          TA            5        Typ
  Fireplaces FireplaceQu GarageType GarageYrBlt GarageFinish GarageCars
1          0        <NA>     Attchd        2003          RFn          2
2          1          TA     Attchd        1976          RFn          2
3          1          TA     Attchd        2001          RFn          2
4          1          Gd     Detchd        1998          Unf          3
5          1          TA     Attchd        2000          RFn          3
6          0        <NA>     Attchd        1993          Unf          2
  GarageArea GarageQual GarageCond PavedDrive WoodDeckSF OpenPorchSF
1        548         TA         TA          Y          0          61
2        460         TA         TA          Y        298           0
3        608         TA         TA          Y          0          42
4        642         TA         TA          Y          0          35
5        836         TA         TA          Y        192          84
6        480         TA         TA          Y         40          30
  EnclosedPorch X3SsnPorch ScreenPorch PoolArea PoolQC Fence MiscFeature
1             0          0           0        0   <NA>  <NA>        <NA>
2             0          0           0        0   <NA>  <NA>        <NA>
3             0          0           0        0   <NA>  <NA>        <NA>
4           272          0           0        0   <NA>  <NA>        <NA>
5             0          0           0        0   <NA>  <NA>        <NA>
6             0        320           0        0   <NA> MnPrv        Shed
  MiscVal MoSold YrSold SaleType SaleCondition SalePrice
1       0      2   2008       WD        Normal    208500
2       0      5   2007       WD        Normal    181500
3       0      9   2008       WD        Normal    223500
4       0      2   2006       WD       Abnorml    140000
5       0     12   2008       WD        Normal    250000
6     700     10   2009       WD        Normal    143000
nb_row <- nrow(df)
nb_col <- ncol(df)
if (nb_row == 1460)
  print("Good")
if (nb_col == 81)
> nb_row <- nrow(df)
> nb_col <- ncol(df)
> if (nb_row == 1460)
+   print("Good")
[1] "Good"
> if (nb_col == 81)
+   print("GOOD2")
[1] "GOOD2"

``` {r}
str(df)
```

> str(df)
'data.frame':   1460 obs. of  81 variables:
 $ Id           : int  1 2 3 4 5 6 7 8 9 10 ...
 $ MSSubClass   : int  60 20 60 70 60 50 20 60 50 190 ...
 $ MSZoning     : chr  "RL" "RL" "RL" "RL" ...
 $ LotFrontage  : int  65 80 68 60 84 85 75 NA 51 50 ...
 $ LotArea      : int  8450 9600 11250 9550 14260 14115 10084 10382 6120 7420 ...
 $ Street       : chr  "Pave" "Pave" "Pave" "Pave" ...
 $ Alley        : chr  NA NA NA NA ...
 $ LotShape     : chr  "Reg" "Reg" "IR1" "IR1" ...
 $ LandContour  : chr  "Lvl" "Lvl" "Lvl" "Lvl" ...
 $ Utilities    : chr  "AllPub" "AllPub" "AllPub" "AllPub" ...
 $ LotConfig    : chr  "Inside" "FR2" "Inside" "Corner" ...
 $ LandSlope    : chr  "Gtl" "Gtl" "Gtl" "Gtl" ...
 $ Neighborhood : chr  "CollgCr" "Veenker" "CollgCr" "Crawfor" ...
 $ Condition1   : chr  "Norm" "Feedr" "Norm" "Norm" ...
 $ Condition2   : chr  "Norm" "Norm" "Norm" "Norm" ...
 $ BldgType     : chr  "1Fam" "1Fam" "1Fam" "1Fam" ...
 $ HouseStyle   : chr  "2Story" "1Story" "2Story" "2Story" ...
 $ OverallQual  : int  7 6 7 7 8 5 8 7 7 5 ...
 $ OverallCond  : int  5 8 5 5 5 5 5 6 5 6 ...
 $ YearBuilt    : int  2003 1976 2001 1915 2000 1993 2004 1973 1931 1939 ...
 $ YearRemodAdd : int  2003 1976 2002 1970 2000 1995 2005 1973 1950 1950 ...
 $ RoofStyle    : chr  "Gable" "Gable" "Gable" "Gable" ...
 $ RoofMatl     : chr  "CompShg" "CompShg" "CompShg" "CompShg" ...
 $ Exterior1st  : chr  "VinylSd" "MetalSd" "VinylSd" "Wd Sdng" ...
 $ Exterior2nd  : chr  "VinylSd" "MetalSd" "VinylSd" "Wd Shng" ...
 $ MasVnrType   : chr  "BrkFace" "None" "BrkFace" "None" ...
 $ MasVnrArea   : int  196 0 162 0 350 0 186 240 0 0 ...
 $ ExterQual    : chr  "Gd" "TA" "Gd" "TA" ...
 $ ExterCond    : chr  "TA" "TA" "TA" "TA" ...
 $ Foundation   : chr  "PConc" "CBlock" "PConc" "BrkTil" ...
 $ BsmtQual     : chr  "Gd" "Gd" "Gd" "TA" ...
 $ BsmtCond     : chr  "TA" "TA" "TA" "Gd" ...
 $ BsmtExposure : chr  "No" "Gd" "Mn" "No" ...
 $ BsmtFinType1 : chr  "GLQ" "ALQ" "GLQ" "ALQ" ...
 $ BsmtFinSF1   : int  706 978 486 216 655 732 1369 859 0 851 ...
 $ BsmtFinType2 : chr  "Unf" "Unf" "Unf" "Unf" ...
 $ BsmtFinSF2   : int  0 0 0 0 0 0 0 32 0 0 ...
 $ BsmtUnfSF    : int  150 284 434 540 490 64 317 216 952 140 ...
 $ TotalBsmtSF  : int  856 1262 920 756 1145 796 1686 1107 952 991 ...
 $ Heating      : chr  "GasA" "GasA" "GasA" "GasA" ...
 $ HeatingQC    : chr  "Ex" "Ex" "Ex" "Gd" ...
 $ CentralAir   : chr  "Y" "Y" "Y" "Y" ...
 $ Electrical   : chr  "SBrkr" "SBrkr" "SBrkr" "SBrkr" ...
 $ X1stFlrSF    : int  856 1262 920 961 1145 796 1694 1107 1022 1077 ...
 $ X2ndFlrSF    : int  854 0 866 756 1053 566 0 983 752 0 ...
 $ LowQualFinSF : int  0 0 0 0 0 0 0 0 0 0 ...
 $ GrLivArea    : int  1710 1262 1786 1717 2198 1362 1694 2090 1774 1077 ...
 $ BsmtFullBath : int  1 0 1 1 1 1 1 1 0 1 ...
 $ BsmtHalfBath : int  0 1 0 0 0 0 0 0 0 0 ...
 $ FullBath     : int  2 2 2 1 2 1 2 2 2 1 ...
 $ HalfBath     : int  1 0 1 0 1 1 0 1 0 0 ...
 $ BedroomAbvGr : int  3 3 3 3 4 1 3 3 2 2 ...
 $ KitchenAbvGr : int  1 1 1 1 1 1 1 1 2 2 ...
 $ KitchenQual  : chr  "Gd" "TA" "Gd" "Gd" ...
 $ TotRmsAbvGrd : int  8 6 6 7 9 5 7 7 8 5 ...
 $ Functional   : chr  "Typ" "Typ" "Typ" "Typ" ...
 $ Fireplaces   : int  0 1 1 1 1 0 1 2 2 2 ...
 $ FireplaceQu  : chr  NA "TA" "TA" "Gd" ...
 $ GarageType   : chr  "Attchd" "Attchd" "Attchd" "Detchd" ...
 $ GarageYrBlt  : int  2003 1976 2001 1998 2000 1993 2004 1973 1931 1939 ...
 $ GarageFinish : chr  "RFn" "RFn" "RFn" "Unf" ...
 $ GarageCars   : int  2 2 2 3 3 2 2 2 2 1 ...
 $ GarageArea   : int  548 460 608 642 836 480 636 484 468 205 ...
 $ GarageQual   : chr  "TA" "TA" "TA" "TA" ...
 $ GarageCond   : chr  "TA" "TA" "TA" "TA" ...
 $ PavedDrive   : chr  "Y" "Y" "Y" "Y" ...
 $ WoodDeckSF   : int  0 298 0 0 192 40 255 235 90 0 ...
 $ OpenPorchSF  : int  61 0 42 35 84 30 57 204 0 4 ...
 $ EnclosedPorch: int  0 0 0 272 0 0 0 228 205 0 ...
 $ X3SsnPorch   : int  0 0 0 0 0 320 0 0 0 0 ...
 $ ScreenPorch  : int  0 0 0 0 0 0 0 0 0 0 ...
 $ PoolArea     : int  0 0 0 0 0 0 0 0 0 0 ...
 $ PoolQC       : chr  NA NA NA NA ...
 $ Fence        : chr  NA NA NA NA ...
 $ MiscFeature  : chr  NA NA NA NA ...
 $ MiscVal      : int  0 0 0 0 0 700 0 350 0 0 ...
 $ MoSold       : int  2 5 9 2 12 10 8 11 4 1 ...
 $ YrSold       : int  2008 2007 2008 2006 2008 2009 2007 2009 2008 2008 ...
 $ SaleType     : chr  "WD" "WD" "WD" "WD" ...
 $ SaleCondition: chr  "Normal" "Normal" "Normal" "Abnorml" ...
 $ SalePrice    : int  208500 181500 223500 140000 250000 143000 307000 200000 129900 118000 ...


 ## 2

``` {r}
print(paste(mean(df$SalePrice)))
hist(df$SalePrice, breaks = 52)
abline(v=mean(df$SalePrice),col="blue",lwd=2)
```


```{r}
print(paste(mean(df$SalePrice)))
print(paste(median(df$SalePrice)))
print(paste(sd(df$SalePrice)))
```

> print(paste(mean(df$SalePrice)))
[1] "180921.195890411"
print(paste(median(df$SalePrice)))
> print(paste(median(df$SalePrice)))
[1] "163000"
print(paste(sd(df$SalePrice)))
> print(paste(sd(df$SalePrice)))
[1] "79442.5028828866"

```{r}
hist(df$TotRmsAbvGrd, breaks = 52)
abline(v=mean(df$TotRmsAbvGrd),col="blue",lwd=2)
```

```{r}
print(paste(mean(df$TotRmsAbvGrd)))
print(paste(median(df$TotRmsAbvGrd)))
print(paste(sd(df$TotRmsAbvGrd)))
```

print(paste(mean(df$TotRmsAbvGrd)))
[1] "6.51780821917808"
print(paste(median(df$TotRmsAbvGrd)))
print(paste(sd(df$TotRmsAbvGrd)))
> print(paste(median(df$TotRmsAbvGrd)))
[1] "6"
> print(paste(sd(df$TotRmsAbvGrd)))
[1] "1.62539329058405"

```{r}
hist(df$OverallCond, breaks = 52)
abline(v=mean(df$OverallCond),col="blue",lwd=2)
```


``` {r}
print(paste(mean(df$OverallCond)))
print(paste(median(df$OverallCond)))
print(paste(sd(df$OverallCond)))
```

print(paste(mean(df$OverallCond)))
[1] "5.57534246575342"
print(paste(median(df$OverallCond)))
> print(paste(median(df$OverallCond)))
[1] "5"
print(paste(sd(df$OverallCond)))
> print(paste(sd(df$OverallCond)))
[1] "1.11279933671273"


## 3

```{r}
index <- grep("SalePrice", colnames(df))
below_average_condition <- df[df$OverallCond < 5, index]
average_condition <- df[df$OverallCond == 5, index]
above_average_condition <- df[df$OverallCond > 5, index]
```


```{r}
print(below_average_condition)
nb1 <- length(seq(min(df$SalePrice),max(df$SalePrice),by=median(df$SalePrice)/20))
hist(average_condition, col='blue', breaks = nb1, xlim = c(0,750000), ylim = c(0,70))
hist(above_average_condition, col='green', add=TRUE, breaks = nb1, xlim = c(0,750000), ylim = c(0,70))
hist(below_average_condition, col='red', add=TRUE, breaks = nb1, xlim = c(0,750000), ylim = c(0,70))
```

## 4

```{r}
idx <- 0
tmp <- 0
corr <- 0
IDX <- 0
for(i in 1:ncol(df)-1) {
  if (is.numeric(df[1,i])) {
    tmp <- cor(df[,i], df$SalePrice)
    if ((corr < tmp) & (!is.na(tmp))) {
      corr <- cor(df[,i], df$SalePrice)
      IDX <- idx
    }
  }
  idx <- idx + 1
}
print(paste("Most Positively Correlated Column:", names(df)[IDX]))
print(paste("Maximum Correlation Value:", corr))
```

> print(paste("Most Positively Correlated Column:", names(df)[IDX]))
[1] "Most Positively Correlated Column: OverallQual"
> print(paste("Maximum Correlation Value:", corr))
[1] "Maximum Correlation Value: 0.790981600583805"

```{r}
idx <- 0
tmp <- 0
corr <- 1
IDX <- 0
for(i in 1:ncol(df)-1) {
  if (is.numeric(df[1,i])) {
    tmp <- cor(df[,i], df$SalePrice)
    if ((corr > tmp) & (!is.na(tmp))) {
      corr <- cor(df[,i], df$SalePrice)
      IDX <- idx
    }
  }
  idx <- idx + 1
}
print(paste("Most Negatively Correlated Column:", names(df)[IDX]))
print(paste("Minimum Correlation Value:", corr))
```

> print(paste("Most Negatively Correlated Column:", names(df)[IDX]))
[1] "Most Negatively Correlated Column: KitchenAbvGr"
> print(paste("Minimum Correlation Value:", corr))
[1] "Minimum Correlation Value: -0.135907370842141"